---
title: "Predicting Police Fatalities in US Metropolis"
author: "Kevin Pham"
date: "June 13, 2019"
output: html_document
---

Libraries
```{r}
library(readr)
library(tidyr)
library(dplyr)
library(lubridate)
library(stringr)

```



```{r, message=FALSE, warning=FALSE}
train <- read_csv("Police shootings/train.csv")
test <- read_csv("Police shootings/test.csv")

sapply(train, function(x) sum(is.na(x))) # count of all NAs in each column

# Getting rid of data that has "U" in Fatal
train <- train[-which(train$Fatal == "U"),]

# # Adding "Fatality" column to testing set with NA values
# test$Fatal <- NA
# 
# # Cleaning training and testing data at the same time
# alldata <- rbind(test,train)

```



```{r, message=FALSE, warning=FALSE}
# We want to get rid of unreliable "1905" dates
# train <- train %>% mutate(year = as.numeric(format(Date, format = "%Y")),
#                           month = as.numeric(format(Date, format = "%m")),
#                           day = as.numeric(format(Date, format = "%d"))
#                           )


train <- separate(data=train, col=Date, sep='/', into = c("month", "day", "year"))
train <- train[-which(train$year == 1905),]

```



```{r}
# Investigating NumberOfOfficers
max(train$NumberOfOfficers, na.rm = TRUE)
table(train$NumberOfOfficers)
sum(prop.table(table(train$NumberOfOfficers))[2:4])
# About 93% of all police reports in our training set have 1, 2, or 3 officers present at the incident, whether or not a fatality took place. My next question is, should I focus my attention on those incidents where 1-3 officers are present at the incident?

df <- data.frame(numOfficers = as.vector(table(train$NumberOfOfficers, train$Fatal)[,1]), Fatalities = as.vector(table(train$NumberOfOfficers, train$Fatal)[,2])) 
freq_df <- data.frame(index = 1:length(df[,1]), numOfficers = df[,1]/sum(df[,1]), Fatalities = df[,2]/sum(df[,2]))
sum(freq_df[2:7,"numOfficers"])

```
# About 98% of all fatalities in the training set take place when 1-6 officers are present. For now, we will only focus on incidents where less than 7 officers were present at the incident. 



```{r}
dim(train) # 2530 obvs
train <- train[which(train$NumberOfOfficers < 7 & train$NumberOfOfficers > 0), ]
dim(train) # 2375 obvs


# Cleaning OfficerRace, OfficerGender
table(train$OfficerRace)
train$OfficerRace <- str_replace_all(train$OfficerRace, c("BLACK" = "B", "WHITE" = "W", "Unknown" = "U", "ASIAN" = "A"))
train$OfficerRace <- str_replace(train$OfficerRace, " ", "")
train$OfficerRace1 <- train$OfficerRace # I want to retain OfficerRace in case we need it
train <- separate(data = train, col = OfficerRace1, sep = ';', into = c("O1Race", "O2Race", "O3Race", "O4Race", "O5Race", "O6Race"), extra = "drop")

train$OfficerGender <- str_replace_all(train$OfficerGender, c("Unknown" = "U"))
train$OfficerGender1 <- train$OfficerGender
train <- separate(data=train, OfficerGender1, sep = ";", into = c("O1Gender", "O2Gender", "O3Gender", "O4Gender", "O5Gender", "O6Gender"), extra = "drop")
```




```{r}
# Cleaning subject's age
table(train$SubjectAge) #oh god
train$SubjectAge <- str_replace_all(train$SubjectAge, c("Juvenile" = '0-19', "N/A" = 'U', "UNKNOWN" = 'U'))

train <- train %>% mutate(SubjectAge = case_when(
  SubjectAge >= '0' & SubjectAge <= '19.0' & SubjectAge != "U" ~ '0-19',
  SubjectAge > '19' & SubjectAge <= '29.0' & SubjectAge != "U" ~ '20-29',
  SubjectAge > '29' & SubjectAge <= '39.0' & SubjectAge != "U" ~ '30-39',
  SubjectAge > '39' & SubjectAge <= '49.0' & SubjectAge != "U" ~ '40-49',
  SubjectAge > '49' & SubjectAge <= '59.0' & SubjectAge != "U" ~ '50-59',
  SubjectAge > '59' & SubjectAge != "U" ~ '60+',
  TRUE ~ "U"
))
```


```{r}
# Removing NAs
# Cleaning SubjectArmed
sum(is.na(train$SubjectArmed)) 
train$SubjectArmed[which(is.na(train$SubjectArmed))] <- 'U'
table(train$SubjectArmed)

# Cleaning SubjectRace
sum(is.na(train$SubjectRace))
train$SubjectRace[which(is.na(train$SubjectRace))] <- 'U'
table(train$SubjectRace)

# Cleaning OfficerRace
sum(is.na(train$OfficerRace))
sum(is.na(train$O1Race))
train$OfficerRace[which(is.na(train$OfficerRace))] <- 'U'
train$O1Race[which(is.na(train$O1Race))] <- 'U'
```



NatureOfStop, NumberOfShots, NumberOfOfficers, Department, FullNarrative, City, Notes
```{r}
train$NatureOfStop <- toupper(train$NatureOfStop)
train$Notes <- toupper(train$Notes)
train$FullNarrative <- toupper(train$FullNarrative)
# table(train$NatureOfStop[-which(is.na(train$NatureOfStop))])
                         
train %>% select(Fatal, SubjectArmed, NatureOfStop, Notes) %>% filter(str_detect(NatureOfStop, "MENTAL")) # most nature of stops that involve mental health usually ends in a fatality

train %>% select(Fatal, SubjectArmed, NatureOfStop, Notes) %>% filter(str_detect(NatureOfStop, "ATTEMPT")) # sometimes, there are cases of "suicide by cop", but there's not enough information to tell us when that takes place, if at all


# defining different types of crime and misc.
drug_related <- c("DRUG", "NARCOTIC", "SUBSTANCE")
property_crime <- c("BURGLARY", "LARCENY", "THEFT", "SHOPLIFT", "STOLE", "THIEF")
nonviolence <- c("SUICIDE", "SUICIDAL", "MENTAL", "TRAFFIC", "CALL", "WARRANT", "SEARCH", "STOP", "SUSPICIOUS")
violence <- c("HOMICIDE", "RAPE", "SEXUAL ASSAULT", "ROB", "ASSAULT",
              "STALKING", "INTIMIDATION", "BATTERY", "SHOOT", "CARJACK",
              "STAB", "KIDNAP", "HOSTAGE", "STABBED")
car_involved <- c("CAR", "VEHICLE", "CHASE")
domestic_related <- c("DOMESTIC", "INVASION", "HOUSE", "HOME")
nonviolence <- c("TRAFFIC")
suspicious_nature <- c("SUSPICIOUS")
armed_crime <- c("ARMED", "WEAPON", "FIREARM")
mental_health <- c("MENTAL", "CRISIS", "EMOTIONAL", "SUICID")
suicide_attempt <- c("SUICIDE", "ATTEMPT", "SUICIDAL")
phone_call <- c("CALL FOR SERVICE", "CALL")
thief_related <- c("THIEF", "STOLEN", "STOLE", "STEAL", "ROBBERY", "BURGLARY", "SHOPLIFT", "THEFT")


# number of observations
n_train <- nrow(train)

# function to paste according to str_detect() format
clean_paste <- function(x) {paste(x, collapse = "|")}

# inputting TRUE/FALSE values for strings we want to detect
train$drug_related <- rep(FALSE, n_train)
train$drug_related[which(str_detect(train$NatureOfStop, clean_paste(drug_related)))] <- TRUE

train$property_crime <- rep(FALSE, n_train)
train$property_crime[which(str_detect(train$NatureOfStop, clean_paste(property_crime)))] <- TRUE

train$violence <- rep(FALSE, n_train)
train$violence[which(str_detect(train$NatureOfStop, clean_paste(violence)))] <- TRUE

train$car_involved <- rep(FALSE, n_train)
train$car_involved[which(str_detect(train$NatureOfStop, clean_paste(car_involved)))] <- TRUE

train$domestic_related <- rep(FALSE, n_train)
train$domestic_related[which(str_detect(train$NatureOfStop, clean_paste(domestic_related)))] <- TRUE

train$nonviolence <- rep(FALSE, n_train)
train$nonviolence[which(str_detect(train$NatureOfStop, clean_paste(nonviolence)))] <- TRUE

train$phone_call <- rep(FALSE, n_train)
train$phone_call[which(str_detect(train$NatureOfStop, clean_paste(phone_call)))] <- TRUE

train$suspicious_nature <- rep(FALSE, n_train)
train$suspicious_nature[which(str_detect(train$NatureOfStop, clean_paste(suspicious_nature)))] <- TRUE

train$armed_crime <- rep(FALSE, n_train)
train$armed_crime[which(str_detect(train$NatureOfStop, clean_paste(armed_crime)))] <- TRUE

train$mental_health <- rep(FALSE, n_train)
train$mental_health[which(str_detect(train$NatureOfStop, clean_paste(mental_health)))] <- TRUE

train$suicide_attempt <- rep(FALSE, n_train)
train$suicide_attempt[which(str_detect(train$NatureOfStop, clean_paste(suicide_attempt)))] <- TRUE

train$thief_related <- rep(FALSE, n_train)
train$thief_related[which(str_detect(train$NatureOfStop, clean_paste(thief_related)))] <- TRUE

train$nonviolence <- rep(FALSE, n_train)
train$nonviolence[which(str_detect(train$NatureOfStop, clean_paste(nonviolence)))] <- TRUE


```
Looking at BJS' website, I classify NatureOfStop as either drug_related, property_crime, violence, or nonviolence. After further analysis, I also created variables car_involved (whether or not there was a car involved in NatureOfStop), domestic_related, suspicious_nature (if a subject is reported as "SUSPICIOUS"), armed_crime (if they were carrying a firearm), mental health (if NatureOfStop involves mental health/mental crisis) and suicide_attempt (if NatureOfStop involved a suicide attempt, which usually resulted in a fatality), phone_call (if a phone call was involved)




Notes (very messy)
```{r, message=FALSE, warning=FALSE}
train$Notes <- toupper(train$Notes)
no_hits <- c("NO HIT", "NON-HIT", "NO-HIT", "NON-FATAL", "NON FATAL", "NONFATAL")
train$no_hits <- rep(FALSE, n_train)
train$no_hits[which(str_detect(train$Notes, clean_paste(no_hits)))] <- TRUE

# train$bbgun <- rep(FALSE, 2375)
# train$bbgun[which(str_detect(train$Notes, "BB GUN"))] <- TRUE
```



FullNarrative (even messier)
```{r}
train$FullNarrative <- toupper(train$FullNarrative)
death_in_narrative <- c("FATAL", "DIED", "DEAD", "DIE", "DOA", "VICTIM", "CASUALTY", "MORTAL", "DECEASE", "KILL", "DEATH", "DECEASE")
train$death_in_narrative <- rep(FALSE, n_train)
train$death_in_narrative[which(str_detect(train$FullNarrative, clean_paste(death_in_narrative)))] <- TRUE


```



NumberOfShots
```{r}
train <- train %>% mutate(Shots = case_when(
  NumberOfShots == 'not clear' | NumberOfShots == 'Unknown' | NumberOfShots == "no information" | NumberOfShots == "U" ~ 'U',
  NumberOfShots == "1*" | NumberOfShots == "1" | NumberOfShots == ">/=1" | NumberOfShots == "2" | NumberOfShots == "0;2;0" 
    |   NumberOfShots == "1;1" | NumberOfShots == ">1" | NumberOfShots == "Multiple" | NumberOfShots == ">/=2" 
    | NumberOfShots == "3" | NumberOfShots == ">2" | NumberOfShots == ">/=3" | NumberOfShots == "4" 
    | NumberOfShots == ">/=4" | NumberOfShots == "5" | NumberOfShots == ">4" | NumberOfShots == "2;3" ~ "1-5",
  NumberOfShots == ">/=5" | NumberOfShots == "6" |NumberOfShots == "7" | NumberOfShots == "8" | NumberOfShots == "9" 
    | NumberOfShots == "10" | NumberOfShots == "5;4" ~ "6-10",
  NumberOfShots == ">/=12" | NumberOfShots == "11" | NumberOfShots == "12" | NumberOfShots == "13" 
    | NumberOfShots == "14" | NumberOfShots == "15" | NumberOfShots == "8;5" 
    | NumberOfShots == ">/=12" ~ "11-15",
  NumberOfShots == "16" | NumberOfShots == "17" | NumberOfShots == "18" | NumberOfShots == "19" 
    | NumberOfShots == "20" | NumberOfShots == "9;9;1" | NumberOfShots == "13;4"~ "16-20",
  NumberOfShots == ">/=20" | NumberOfShots == "21" | NumberOfShots == "22" | NumberOfShots == "23" 
    | NumberOfShots == "24" | NumberOfShots == "25" | NumberOfShots == "26" | NumberOfShots == "27" 
    | NumberOfShots == "28" | NumberOfShots == "29" | NumberOfShots == "30" | NumberOfShots == "5;8;6;4;6" 
    | NumberOfShots == "13;9" | NumberOfShots == "16;2;1;2" ~ "21-30",
  NumberOfShots == "31" | NumberOfShots == "34" | NumberOfShots == "4;14;16" | NumberOfShots == "40" 
    | NumberOfShots == "42859" | NumberOfShots == "43" | NumberOfShots == "45" | NumberOfShots == "46" 
    | NumberOfShots == "52" | NumberOfShots == "62" | NumberOfShots == "65" | NumberOfShots == "72" 
    | NumberOfShots == "16;16;2;12" ~ "30+",
  TRUE ~ NA_character_
)) 


```




City, Department
```{r}
# mappingpoliceviolence.org 
top25_departments <- c("St. Louis", "Phoenix", "Kansas City", "Columbus", "Albuquerque", "Baltimore", "Tucson", "Jacksonville")
top10_police <- c("ALBUQUERQUE", "LASVEGAS", "WASHINGTON DC", "PHOENIX", "TUCSON", "LOSANGELES", "SANJOSE", "SAN FRANCISCO", "SANDIEGO", "FAIRFAXCOUNTY")
top25_poverty <- c("COLUMBUS", "ALBUQUERQUE", "NEWORLEANS", "WASHINGTON DC", "LOUISVILLE", "WASHINGTON DC", "NASHVILLE", "MEMPHIS", "ATLANTA", "PHOENIX", "TUCSON", "HOUSTON", "DALLAS", "AUSTIN", "SAN ANTONION", "EL PASO", "FORT WORTH", "NEW YORK", "MIAMI", "JACKSONVILLE", "INDIANAPOLIS", "ST. LOUIS", "KANSAS CITY", "LOSANGELES", "SANJOSE", "SAN FRANCISCO", "SANDIEGO", "FAIRFAXCOUNTY", "PORTLAND", "LASVEGAS")


# rates of police killings by state, 2013-2018
train$top_state <- rep(FALSE, n_train)

# police departments bearing most responsibility for police violence, 2013-2018
train$top25_departments <- rep(FALSE, n_train)
train$top25_departments[which(str_detect(train$Department, clean_paste(top25_departments)))] <- TRUE

# top10_police states with highest rates of police violence
train$top10_police <- rep(FALSE, n_train)
train %>% select(City, Department) %>% filter(str_detect(train$City, "WASHINGTON DC"))
train$top10_police[which(str_detect(train$City,clean_paste(top10_states)))] <- TRUE

# us states by poverty rate, 2017
# source: wiki
train$top25_poverty <- rep(FALSE, n_train)

train$top25_poverty[which(str_detect(train$City, clean_paste(top25_poverty)))] <- TRUE
```




Type Conversion
```{r, warning=T}
  names(train)
  train$id <- factor(train$id)
  train$NumberOfSubjects <- factor(train$NumberOfSubjects)
  train$SubjectArmed <- factor(train$SubjectArmed)
  train$SubjectRace <- factor(train$SubjectRace)
  train$SubjectGender <- factor(train$SubjectGender)
  train$Age <- factor(train$Age)
  train$NumberOfOfficers <- factor(train$NumberOfOfficers)
  train$OfficerRace <- factor(train$OfficerRace)
  train$OfficerGender <- factor(train$OfficerGender)
  train$O1Gender <- factor(train$O1Gender)
  train$O2Gender <- factor(train$O2Gender)
  train$O3Gender <- factor(train$O3Gender)
  train$O4Gender <- factor(train$O4Gender)
  train$O5Gender <- factor(train$O5Gender)
  train$O6Gender <- factor(train$O6Gender)
  train$O1Race <- factor(train$O1Race)
  train$O2Race <- factor(train$O2Race)
  train$O3Race <- factor(train$O3Race)
  train$O4Race <- factor(train$O4Race)
  train$O5Race <- factor(train$O5Race)
  train$O6Race <- factor(train$O6Race)
  train$Shots <- factor(train$Shots)
  train$Fatal <- factor(train$Fatal)


```





```{r}
# set.seed(5)
# 
# # randomforest
# library(randomForest)
# m3 <- randomForest(Fatal ~ SubjectArmed + SubjectRace + SubjectGender + Age + O1Race + thief_related + domestic_related + no_hits + death_in_narrative + mental_health + top10_police + top25_poverty + top25_departments + O1Gender + nonviolence,
#                    data = train, mtry = 3, importance = TRUE, na.action = na.exclude)
# print(m3)
# 
# # Inspecting OOB error matrix
# err <- m3$err.rate
# head(err)
# 
# # final OOB error rate (last row in err matrix)
# oob_err <- err[nrow(err), "OOB"]
# print(oob_err)
# 
# 
# importance <- importance(m3)
# varImpPlot(m3)
# 
# # Plotting Model
# plot(m3)
# legend(x = "right", legend = colnames(err), fill = 1:ncol(err))

# # varImpPlot tells us which variables the model believes to be more important when predicting a fatality.
```





PART B) Cleaning Test Data
```{r}
test <- read_csv("Police shootings/test.csv")
test <- test %>% mutate(year = as.numeric(format(Date, format = "%Y")),
                          month = as.numeric(format(Date, format = "%m")),
                          day = as.numeric(format(Date, format = "%d"))
                          )


# test <- separate(data=test, col=Date, sep='/', into = c("month", "day", "year"))



# Cleaning OfficerRace, OfficerGender
test$OfficerRace <- str_replace_all(test$OfficerRace, c("BLACK" = "B", "WHITE" = "W", "Unknown" = "U", "ASIAN" = "A"))
test$OfficerRace <- str_replace(test$OfficerRace, " ", "")
test$OfficerRace1 <- test$OfficerRace # I want to retain OfficerRace in case we need it
test <- separate(data = test, col = OfficerRace1, sep = ';', into = c("O1Race", "O2Race", "O3Race", "O4Race", "O5Race", "O6Race"), extra = "drop")

test$OfficerGender <- str_replace_all(test$OfficerGender, c("Unknown" = "U"))
test$OfficerGender1 <- test$OfficerGender
test <- separate(data=test, OfficerGender1, sep = ";", into = c("O1Gender", "O2Gender", "O3Gender", "O4Gender", "O5Gender", "O6Gender"), extra = "drop")


# Cleaning subject's age
test$SubjectAge <- str_replace_all(test$SubjectAge, c("Juvenile" = '0-19', "N/A" = 'U', "UNKNOWN" = 'U'))

test <- test %>% mutate(Age = case_when(
  SubjectAge >= '0' & SubjectAge <= '19.0' & SubjectAge != "U" ~ '0-19',
  SubjectAge > '19' & SubjectAge <= '29.0' & SubjectAge != "U" ~ '20-29',
  SubjectAge > '29' & SubjectAge <= '39.0' & SubjectAge != "U" ~ '30-39',
  SubjectAge > '39' & SubjectAge <= '49.0' & SubjectAge != "U" ~ '40-49',
  SubjectAge > '49' & SubjectAge <= '59.0' & SubjectAge != "U" ~ '50-59',
  SubjectAge > '59' & SubjectAge != "U" ~ '60+',
  TRUE ~ "U"
))


# Cleaning SubjectArmed
sum(is.na(test$SubjectArmed)) 
test$SubjectArmed[which(is.na(test$SubjectArmed))] <- 'U'

# Cleaning SubjectRace
sum(is.na(test$SubjectRace))
test$SubjectRace[which(is.na(test$SubjectRace))] <- 'U'

# Cleaning OfficerRace
test$OfficerRace[which(is.na(test$OfficerRace))] <- 'U'
test$O1Race[which(is.na(test$O1Race))] <- 'U'


# Cleaning NatureOfStop
test$NatureOfStop <- toupper(test$NatureOfStop)
test$Notes <- toupper(test$Notes)
test$FullNarrative <- toupper(test$FullNarrative)
                       

# defining different types of crime and misc.
drug_related <- c("DRUG", "NARCOTIC", "SUBSTANCE")
property_crime <- c("BURGLARY", "LARCENY", "THEFT", "SHOPLIFT", "STOLE", "THIEF")
nonviolence <- c("SUICIDE", "SUICIDAL", "MENTAL", "TRAFFIC", "CALL", "WARRANT", "SEARCH", "STOP", "SUSPICIOUS")
violence <- c("HOMICIDE", "RAPE", "SEXUAL ASSAULT", "ROB", "ASSAULT",
              "STALKING", "INTIMIDATION", "BATTERY", "SHOOT", "CARJACK",
              "STAB", "KIDNAP", "HOSTAGE", "STABBED")
car_involved <- c("CAR", "VEHICLE", "CHASE")
domestic_related <- c("DOMESTIC", "INVASION", "HOUSE", "HOME")
nonviolence <- c("TRAFFIC")
suspicious_nature <- c("SUSPICIOUS")
armed_crime <- c("ARMED", "WEAPON", "FIREARM")
mental_health <- c("MENTAL", "CRISIS", "EMOTIONAL", "SUICID")
suicide_attempt <- c("SUICIDE", "ATTEMPT", "SUICIDAL")
phone_call <- c("CALL FOR SERVICE", "CALL")
thief_related <- c("THIEF", "STOLEN", "STOLE", "STEAL", "ROBBERY", "BURGLARY", "SHOPLIFT", "THEFT")


# number of observations
n_test <- nrow(test)

# function to paste according to str_detect() format
clean_paste <- function(x) {paste(x, collapse = "|")}

# detecting different types of crimes from "Notes"
test$drug_related <- rep(FALSE, n_test)
test$drug_related[which(str_detect(test$NatureOfStop, clean_paste(drug_related)))] <- TRUE

test$property_crime <- rep(FALSE, n_test)
test$property_crime[which(str_detect(test$NatureOfStop, clean_paste(property_crime)))] <- TRUE

test$violence <- rep(FALSE, n_test)
test$violence[which(str_detect(test$NatureOfStop, clean_paste(violence)))] <- TRUE

test$car_involved <- rep(FALSE, n_test)
test$car_involved[which(str_detect(test$NatureOfStop, clean_paste(car_involved)))] <- TRUE

test$domestic_related <- rep(FALSE, n_test)
test$domestic_related[which(str_detect(test$NatureOfStop, clean_paste(domestic_related)))] <- TRUE

test$nonviolence <- rep(FALSE, n_test)
test$nonviolence[which(str_detect(test$NatureOfStop, clean_paste(nonviolence)))] <- TRUE

test$phone_call <- rep(FALSE, n_test)
test$phone_call[which(str_detect(test$NatureOfStop, clean_paste(phone_call)))] <- TRUE

test$suspicious_nature <- rep(FALSE, n_test)
test$suspicious_nature[which(str_detect(test$NatureOfStop, clean_paste(suspicious_nature)))] <- TRUE

test$armed_crime <- rep(FALSE, n_test)
test$armed_crime[which(str_detect(test$NatureOfStop, clean_paste(armed_crime)))] <- TRUE

test$mental_health <- rep(FALSE, n_test)
test$mental_health[which(str_detect(test$NatureOfStop, clean_paste(mental_health)))] <- TRUE

test$suicide_attempt <- rep(FALSE, n_test)
test$suicide_attempt[which(str_detect(test$NatureOfStop, clean_paste(suicide_attempt)))] <- TRUE

test$suspicious_nature <- rep(FALSE, n_test)
test$suspicious_nature[which(str_detect(test$NatureOfStop, clean_paste(suspicious_nature)))] <- TRUE




# Cleaning Notes by finding "no hits"
test$Notes <- toupper(test$Notes)
no_hits <- c("NO HIT", "NON-HIT", "NO-HIT", "NON-FATAL", "NON FATAL", "NONFATAL")
test$no_hits <- rep(FALSE, n_test)
test$no_hits[which(str_detect(test$Notes, clean_paste(no_hits)))] <- TRUE

# Cleaning FullNarrative by finding fatalities
test$FullNarrative <- toupper(test$FullNarrative)
death_in_narrative <- c("FATAL", "DIED", "DEAD", "DIE", "DOA", "VICTIM", "CASUALTY", "MORTAL", "DECEASE", "KILL", "DEATH", "DECEASE")
test$death_in_narrative <- rep(FALSE, n_test)
test$death_in_narrative[which(str_detect(test$FullNarrative, clean_paste(death_in_narrative)))] <- TRUE






# mappingpoliceviolence.org 
top25_departments <- c("St. Louis", "Phoenix", "Kansas City", "Columbus", "Albuquerque", "Baltimore", "Tucson", "Jacksonville")
top10_police <- c("ALBUQUERQUE", "LASVEGAS", "WASHINGTON DC", "PHOENIX", "TUCSON", "LOSANGELES", "SANJOSE", "SAN FRANCISCO", "SANDIEGO", "FAIRFAXCOUNTY")
top25_poverty <- c("COLUMBUS", "ALBUQUERQUE", "NEWORLEANS", "WASHINGTON DC", "LOUISVILLE", "WASHINGTON DC", "NASHVILLE", "MEMPHIS", "ATLANTA", "PHOENIX", "TUCSON", "HOUSTON", "DALLAS", "AUSTIN", "SAN ANTONION", "EL PASO", "FORT WORTH", "NEW YORK", "MIAMI", "JACKSONVILLE", "INDIANAPOLIS", "ST. LOUIS", "KANSAS CITY", "LOSANGELES", "SANJOSE", "SAN FRANCISCO", "SANDIEGO", "FAIRFAXCOUNTY", "PORTLAND", "LASVEGAS")

# top cities for police violence
test$top_city <- rep(FALSE, n_test)
test$top_city[which(str_detect(test$City, clean_paste(top_city)))] <- TRUE

# rates of police killings by state, 2013-2018
test$top_state <- rep(FALSE, n_test)

# police departments bearing most responsibility for police violence, 2013-2018
test$top25_departments <- rep(FALSE, n_test)
test$top25_departments[which(str_detect(test$Department, clean_paste(top25_departments)))] <- TRUE

# top10_police states with highest rates of police violence
test$top10_police <- rep(FALSE, n_test)

test$top10_police[which(str_detect(test$City,clean_paste(top10_states)))] <- TRUE

# us states by poverty rate, 2017
# source: wiki
test$top25_poverty <- rep(FALSE, n_test)
test$top25_poverty[which(str_detect(test$City, clean_paste(top25_poverty)))] <- TRUE

# Cleaning NumberOfShots
test <- test %>% mutate(Shots = case_when(
  NumberOfShots == 'not clear' | NumberOfShots == 'Unknown' | NumberOfShots == "no information" | NumberOfShots == "U" ~ 'U',
  NumberOfShots == "1*" | NumberOfShots == "1" | NumberOfShots == ">/=1" | NumberOfShots == "2" | NumberOfShots == "0;2;0" 
    |   NumberOfShots == "1;1" | NumberOfShots == ">1" | NumberOfShots == "Multiple" | NumberOfShots == ">/=2" 
    | NumberOfShots == "3" | NumberOfShots == ">2" | NumberOfShots == ">/=3" | NumberOfShots == "4" 
    | NumberOfShots == ">/=4" | NumberOfShots == "5" | NumberOfShots == ">4" | NumberOfShots == "2;3" ~ "1-5",
  NumberOfShots == ">/=5" | NumberOfShots == "6" |NumberOfShots == "7" | NumberOfShots == "8" | NumberOfShots == "9" 
    | NumberOfShots == "10" | NumberOfShots == "5;4" ~ "6-10",
  NumberOfShots == ">/=12" | NumberOfShots == "11" | NumberOfShots == "12" | NumberOfShots == "13" 
    | NumberOfShots == "14" | NumberOfShots == "15" | NumberOfShots == "8;5" 
    | NumberOfShots == ">/=12" ~ "11-15",
  NumberOfShots == "16" | NumberOfShots == "17" | NumberOfShots == "18" | NumberOfShots == "19" 
    | NumberOfShots == "20" | NumberOfShots == "9;9;1" | NumberOfShots == "13;4"~ "16-20",
  NumberOfShots == ">/=20" | NumberOfShots == "21" | NumberOfShots == "22" | NumberOfShots == "23" 
    | NumberOfShots == "24" | NumberOfShots == "25" | NumberOfShots == "26" | NumberOfShots == "27" 
    | NumberOfShots == "28" | NumberOfShots == "29" | NumberOfShots == "30" | NumberOfShots == "5;8;6;4;6" 
    | NumberOfShots == "13;9" | NumberOfShots == "16;2;1;2" ~ "21-30",
  NumberOfShots == "31" | NumberOfShots == "34" | NumberOfShots == "4;14;16" | NumberOfShots == "40" 
    | NumberOfShots == "42859" | NumberOfShots == "43" | NumberOfShots == "45" | NumberOfShots == "46" 
    | NumberOfShots == "52" | NumberOfShots == "62" | NumberOfShots == "65" | NumberOfShots == "72" 
    | NumberOfShots == "16;16;2;12" ~ "30+",
  TRUE ~ NA_character_
)) 



  test$id <- factor(test$id)
  test$NumberOfSubjects <- factor(test$NumberOfSubjects)
  test$SubjectArmed <- factor(test$SubjectArmed)
  test$SubjectRace <- factor(test$SubjectRace)
  test$SubjectGender <- factor(test$SubjectGender)
  test$Age <- factor(test$Age)
  test$NumberOfOfficers <- factor(test$NumberOfOfficers)
  test$OfficerRace <- factor(test$OfficerRace)
  test$OfficerGender <- factor(test$OfficerGender)
  test$O1Race <- factor(test$O1Race)
  test$O2Race <- factor(test$O2Race)
  test$O3Race <- factor(test$O3Race)
  test$O4Race <- factor(test$O4Race)
  test$O5Race <- factor(test$O5Race)
  test$O6Race <- factor(test$O6Race)
  test$O1Gender <- factor(test$O1Gender)
  test$O2Gender <- factor(test$O2Gender)
  test$Shots <- factor(test$Shots)
  
  

  
  

```




PART C) TESTING
```{r}
set.seed(5)

library(tree)
m2 <- tree(Fatal ~ SubjectArmed + SubjectRace + SubjectGender + Age + O1Race + thief_related + domestic_related + no_hits + death_in_narrative + mental_health + top10_police + top25_poverty + top25_departments + O1Gender + nonviolence, 
                   data = train, mindev = 0.001)
summary(m2)


#making predictions
probs <- predict(m1, newdata = test, type = "class")
probs <- as.character(probs)
probs[probs == "N"] <- "No"
probs[probs == "F"] <- "Yes"


probs <- predict(m2, newdata = test, type = "class")
probs <- as.character(probs)
probs[probs == "N"] <- "No"
probs[probs == "F"] <- "Yes"
submission <- data.frame(test$id, probs)
names(submission) <- c("id", "Fatal")


write.csv(submission, file = "submission.csv", row.names = FALSE, col.names = c("id", "Fatal"))
```









